---
title: "Smooth Stimson Series ADL Models"
subtitle: "Media Stance"
author: "Suzie and Nitheesha"
date: "July 24, 2023"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_notebook:
    number_sections: no
    theme: cerulean
    toc: yes
    toc_depth: 3
  word_document: null
---
<style type="text/css">

body{ /* Normal  */
      font-size: 14px;
  }
td {  /* Table  */
  font-size: 8px;
}

code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

---

```{r setup, include=FALSE}
library(forecast) 
library(astsa) 
library(dynlm)
library(lmtest)
library(readr)
library(tseries)
library(tidyr)
library(ggplot2)
library(stargazer)
suppressPackageStartupMessages(source("~/OneDrive - The Pennsylvania State University/MainstreamMediaAnalysis/PGTSadfv2.r"))

data <- read_csv("~/OneDrive - The Pennsylvania State University/MainstreamMediaAnalysis/MonthlyEconomicData_Mar292023.csv")
NYT <- read_csv("~/OneDrive - The Pennsylvania State University/MainstreamMediaAnalysis/NYT/stim.NYT.csv")
USA <- read_csv("~/OneDrive - The Pennsylvania State University/MainstreamMediaAnalysis/USAToday/stim.USA.csv")
```

# ADL Models
## NYT Full Period
```{r}
data.ts <- ts(data, start = c(1959,1), end = c(2022,11), freq = 12)

#NYT_adl <- dynlm(NYT_smth~L(NYT_smth,1:3) +
#               inflation + L(inflation, 1:3) +
#               DUNRATE + L(DUNRATE, 1:3) +
#               real_sp_index_pc + L(real_sp_index_pc, 1:3) +
#                DSPIC96_pc_pca + L(DSPIC96_pc_pca,1:3) +
#                LFPR_pca + L(LFPR_pca, 1:3),
#              data=data.ts)
#summary(NYT_adl)

NYT_adl <- dynlm(NYT_smth~L(NYT_smth,c(1,12)) +
               inflation + L(inflation, 1) +
               DUNRATE + L(DUNRATE, 1) +
               real_sp_index_pc + L(real_sp_index_pc, 1) +
                DSPIC96_pc_pca + L(DSPIC96_pc_pca,1) +
                LFPR_pca + L(LFPR_pca, 1),
              data=data.ts)

#stick with this latest model: lag 12 deals with seasonality
summary(NYT_adl)
```

### BG test for serial correlation
Reject the null of no serial correlation at lag order 24.
Fail to reject the null of no serial correlation at lag orders 12 and 6.
```{r}
bgtest(NYT_adl, order=24, type="Chisq")  
bgtest(NYT_adl, order=12, type="Chisq") 
bgtest(NYT_adl, order=6, type="Chisq") 
```

### Residual Plot
```{r}
nyt_res <- resid(NYT_adl)
plot(nyt_res,
     ylab = "Residuals",
     main = "Residuals for NYT ADL Model")
```

## NYT Overlap Period
```{r}
data.olp.ts <- ts(data[340:780,], start = c(1987,4), end = c(2022,11), freq = 12)

#NYT_olp_sm <- dynlm(NYT_smth~L(NYT_smth,1:3) +  
#              inflation + L(inflation, 1:3)  +
#               DUNRATE + L(DUNRATE, 1:3)+
#               real_sp_index_pc + L(real_sp_index_pc, 1:3)  +
#                DSPIC96_pc_pca + L(DSPIC96_pc_pca,1:3)  +
#                LFPR_pca + L(LFPR_pca, 1:3),
#              data=data.olp.ts)

#summary(NYT_olp_sm)

NYT_olp_sm <- dynlm(NYT_smth~L(NYT_smth,c(1,12)) +  
               inflation + L(inflation, 1)  +
               DUNRATE + L(DUNRATE, 1)+
               real_sp_index_pc + L(real_sp_index_pc, 1)  +
                DSPIC96_pc_pca + L(DSPIC96_pc_pca,1)  +
                LFPR_pca + L(LFPR_pca, 1),
              data=data.olp.ts)

summary(NYT_olp_sm)
```

### BG test for serial correlation
Fail to reject the null of no serial correlation at all lag orders. 
```{r}
bgtest(NYT_olp_sm, order=24, type="Chisq")  
bgtest(NYT_olp_sm, order=12, type="Chisq") 
bgtest(NYT_olp_sm, order=6, type="Chisq") 
```

### Residual Plot
```{r}
nyt_olp_res <- resid(NYT_olp_sm)
plot(nyt_olp_res,
     ylab = "Residuals",
     main = "Residuals for NYT Overlapping Period ADL Model")
```
### check residuals
```{r}
nyt_olp_res 
```

## USA Today Full Period

```{r}
data.us.ts <- ts(data[340:780,], start = c(1987,4), end = c(2022,11), freq = 12)

USA_adl <- dynlm(USA_smth~L(USA_smth,c(1,12)) +
               inflation + L(inflation, 1) +
               DUNRATE + L(DUNRATE, 1) +
               real_sp_index_pc + L(real_sp_index_pc, 1) +
                DSPIC96_pc_pca + L(DSPIC96_pc_pca,1) +
                LFPR_pca + L(LFPR_pca, 1),
              data=data.us.ts)
summary(USA_adl)
```

### BG test for serial correlation
Fail to reject the null of no serial correlation at all lag orders. 
```{r}
bgtest(USA_adl, order=24, type="Chisq")  
bgtest(USA_adl, order=12, type="Chisq") 
bgtest(USA_adl, order=6, type="Chisq") 
```
### Residual Plot
```{r}
usa_res <- resid(USA_adl)
plot(usa_res,
     ylab = "Residuals",
     main = "Residuals for USA Today ADL Model")
```
```{r}
acf(usa_res)
```

```{r}
pacf(usa_res)
```

## Final Model
```{r}
stargazer(NYT_adl, NYT_olp_sm, USA_adl,
          title = "ADL Model Output",
          align = TRUE)
```